---
title: "Regression Models Course Final Project"
author: "Mike Holmquest"
date: "4/22/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Executive Summary
This project will be analyzing mileage data for Motor Trend, a magazine about the automobile industry. We will use a data set of a collection of cars, they are interested in exploring the relationship between a set of variables and miles per gallon (MPG) (outcome). They are particularly interested in the following two questions:

Question 1 - “Is an automatic or manual transmission better for MPG”
Question 2 - "Quantify the MPG difference between automatic and manual transmissions"

## Load data and programs

Download the data and associated programs
```{r Download_data and programs}
library(datasets)
data(mtcars)
library(ggplot2)

```

## Exploratory Data Analysis

The first goal is going to be to take a high level look at the data to see its organization and structure.
```{r Data Exploration}
## What do the variables mean
?mtcars
## Use str to see a compact summary of the data
str(mtcars)
sapply(mtcars, class)
```

From the information on the data set we see that the mtcars dataset is a data frame with 32 observations on 11 (numeric) variables.

mpg	Miles/(US) gallon
cyl	Number of cylinders
disp	Displacement (cu.in.)
hp	Gross horsepower
drat	Rear axle ratio
wt	Weight (1000 lbs)
qsec	1/4 mile time
vs	Engine (0 = V-shaped, 1 = straight)
am	Transmission (0 = automatic, 1 = manual)
gear	Number of forward gears
carb	Number of carburetors

From this information we know that all of the variables are numeric. We can also see that vs(Engine Type) and am(Transmission) are "0" , "1' factor variables. We can also see that cyl (Cylinders) is a 3 level variable with "4", "6", and "8". Let's convert these to factor variables for analysis purposes. 

```{r convert to Factor}
mtcars$vs = as.factor(mtcars$vs)
mtcars$am = as.factor(mtcars$am)
levels(mtcars$am) <- c("Automatic", "Manual")
mtcars$cyl = as.factor(mtcars$cyl)
```
The most important variables for our analysis are mpg - miles per gallon, and am - transmission. We will also use some of the other variables lik cyl - cylinders, and wt - weight to see correlations and/or causations. 

Lets take a quick look at the data to see what kind of trends it might show. 

```{r mpg vs. transmission plot}
plot1 <- ggplot(mtcars, aes(y = mpg, x = am, fill = am))
plot1 = plot1 + geom_violin(colour = "black", size = 2)
plot1 = plot1 + xlab("Type of Transmission") + ylab("MPG(miles per gallon)")
plot1 = plot1 + scale_fill_discrete(name = "Transmission Type", labels=c("Automatic", "Manual"))
plot1
## What are some useful summary statistics
aggregate(mpg ~ am, data=mtcars, mean)
```
Just out of curiosity lets take a look at how weight affects mpg.
```{r mpg vs. weight fill = cylinder}
plot2 <- ggplot(mtcars, aes(wt, mpg, colour = cyl)) + geom_point()
plot2 = plot2 + labs(colour = "Cylinders")
plot2 = plot2 + xlab("Weight in Tons") + ylab("MPG(miles per gallon")
plot2

## What are some useful summary statistics
aggregate(mpg ~ wt, data=mtcars, mean)
```


```{r mpg vs. cyl}
p <- ggplot(mtcars, aes(mpg, cyl, colour = am)) + geom_point()
p + labs(colour = "Transmission Type")
```



## Regression Models

We will do two regression models, as follows:
Reg 1 - MPG ~ AM (MPG dependent on AM) Reg 2 - MPG ~ AM + WT + CYL + DISP + HP (MPG dependent on WT, CYL, DISP, and HP)
```{r}
reg1 <- lm(mpg ~ am, mtcars)
reg2 <- lm(mpg ~ am + wt + cyl + disp + hp, mtcars)
```



